Relevance-based Quantization of Scattering Features for Unsupervised Mining of Environmental Audio

### 2018-08-13 paper/lostanlenRelevanceScattering.tex
typos fixes

The amount of audio data recorded from our sonic environment has grown considerably over the past decades.
In order to measure the effect of human activity and climate change on animal biodiversity~\cite{sueur2015ecoacoustics}, researchers have recently undertaken a massive deployment of acoustic sensors throughout the world~\cite{warren2006urban, NessSST13, stowell13b}.
In addition, recent work has explored acoustic monitoring for characterization of human pleasantness in urban areas~\cite{guyot2005urban, ricciardi2015sound}, as well as the prediction of annoyance due to traffic~\cite{gloaguen}.
Since they bear a strong societal impact and raise many scientific challenges, we believe that these applications are of considerable interest for to signal processing community.

@@

With this aim, we need to design an algorithm for acoustic similarity retrieval, where the audio fragments judged ``most similar'' to a given query recording must be extracted from some larger dataset. To construct such an algorithm, we are required to represent an audio recording in a way that captures its distinctive qualities. A widespread choice of representation is the bag of frames~\cite{aucouturier2007bag}, which describes an auditory scene recording using summary statistics of short-time features. Unfortunately, the bag-of-frames approach only captures the average structure of the scene, so the approach often fails when presented with highly dynamic scenes or those characterised by a few distinct sound events. Furthermore, experiments in cognitive psychology~\cite{dubois2006cognitive} and cognitive neuroscience~\cite{nelken2004processing} suggest that human acoustic perception is highly sensitive to such isolated sound events. We believe that the failure to model such distinct events is one of the reasons why the bag-of-frames representation is insufficient~\cite{lagrange:hal-01082501}.

@@

For our task, one advantage of the scattering transform is not requiring a training step, allowing for a wider range of applications compared to learned features. Indeed, for data mining of previously unheard datasets, the properties of relevant audio structures remains to be defined, leading to an unsupervised setting.

@@

While BoF has largely been superseded by more sophisticated methods for the task of acoustic scene classification~\cite{7100934}, it remains the best-performing model for acoustic scene similarity retrieval.
Recently, this representation was shown to perform comparably to direct averaging of the features for a variety of datasets~\cite{lagrange:hal-01082501}.
This contrasts with the typical morphology of acoustic scenes, a ``skeleton of events on a bed of textures,'' where a few discrete sound events are superimposed upon a stationary acoustic background~\cite{nelken2013}.
Such events are not well-characterised by summarizing short-term features, but are better described by large-scale temporal evolution of auditory scenes.
The latter approach should therefore prove more fruitful in measuring auditory scene similarity.

@@

Many algorithms in pattern recognition, including nearest neighbor classifiers and SVMs, tend to work best when all features follow a standard normal distribution across all training instances~\cite{Hsu2003}.
Yet the distribution of the scattering coefficients is skewed towards larger values. We can reduce this skewness by applying a pointwise concave transformation to all coefficients. In particular, we find that the logarithm performs particularly well in this respect.
Figure~\ref{fig:histograms} shows the distribution of an arbitrarily chosen scattering coefficient over the DCASE 2013 dataset, before and after logarithmic compression.

@@

The experiments in this paper are carried out on the DCASE 2013 dataset~\cite{7100934}.
Although the dataset was constructed for the task of acoustic scene classification, where the goal is to correctly assign the class of a given recording, we can use the same recordings and class labels for the task of acoustic scene similarity retrieval.
The dataset consists of two parts, a public and a private subset, each made up of $100$ acoustic scene recordings sampled at $44100\,\mathrm{Hz}$ and $30$ seconds in duration. The dataset is evenly divided into $10$ acoustic scene classes: \emph{bus}, \emph{busy street}, \emph{office}, \emph{open air market}, \emph{park}, \emph{quiet street}, \emph{restaurant}, \emph{supermarket}, \emph{tube}, and \emph{tube station}.
The recordings were made by three different recordists at a wide variety of locations in the Greater London area over a period of several months.
In order to avoid any correlation between recording conditions and label distribution, all recordings were carried out under moderate weather conditions, at varying times of day and week, and each recordist recorded each scene type. As a result, the dataset enjoys significant intra-class diversity while remaining of manageable size, making it suitable for evaluation of algorithmic design choices~\cite{lagrange:hal-01082501}.

@@

The scaling parameter $q$ of the RBF kernels (see Eq.~\ref{eq:kc}) is set to $10\%$ of the number of data points to cluster.  As the number of Gaussians for the BoF approach, the numbers of clusters $M$ controls the level of abstraction. For each method, unless otherwise stated, the parameter $M$ is set to $8$. It thus allows $8$ different types of observations to be modeled, which seems reasonable given the duration of the scene (30 seconds). Note that this is the only free parameter in the proposed method. However, except for $\emph{RbQ-a}$, the results are not very sensitive to the choice of $M$, as long as it is large enough to characterize the seen. For a numerical demonstration of this, please see the end of the next section.

>>>>

The amount of audio data recorded from our sonic environment has grown considerably over the past decades.
In order to measure the effect of human activity and climate change on animal biodiversity~\cite{sueur2015ecoacoustics}, researchers have recently undertaken a massive deployment of acoustic sensors throughout the world~\cite{warren2006urban, NessSST13, stowell13b}.
In addition, recent work has explored acoustic monitoring for characterization of human pleasantness in urban areas~\cite{guyot2005urban, ricciardi2015sound}, as well as the prediction of annoyance due to traffic~\cite{gloaguen}.
Since they bear a strong societal impact and raise many scientific challenges, we believe that these applications are of considerable interest for signal processing community.

@@

With this aim, we need to design an algorithm for acoustic similarity retrieval, where the audio fragments judged ``most similar'' to a given query recording must be extracted from some larger dataset. To construct such an algorithm, we are required to represent an audio recording in a way that captures its distinctive qualities. A widespread choice of representation is the bag of frames~\cite{aucouturier2007bag}, which describes an auditory scene recording using summary statistics of short-time features. Unfortunately, the bag-of-frames approach only captures the average structure of the scene, so the approach often fails when presented with highly dynamic scenes or those characterised by a few distinct sound events sparsely distributed over time. Furthermore, experiments in cognitive psychology~\cite{dubois2006cognitive} and cognitive neuroscience~\cite{nelken2004processing} suggest that human acoustic perception is highly sensitive to such isolated sound events. We believe that the failure to model such distinct events is one of the reasons why the bag-of-frames representation is insufficient~\cite{lagrange:hal-01082501}.

@@

For our task, one advantage of the scattering transform is that it does not require a training step, allowing for a wider range of applications compared to learned features. Indeed, for data mining of previously unheard datasets, the properties of relevant audio structures remains to be defined, leading to an unsupervised setting.

@@

While BoF has largely been superseded by more sophisticated methods for the task of acoustic scene classification~\cite{7100934}, it remains the best-performing model for acoustic scene similarity retrieval.
Though, this representation was recently shown to perform comparably to direct averaging of the features for a variety of datasets~\cite{lagrange:hal-01082501}.
This contrasts with the typical morphology of acoustic scenes, a ``skeleton of events on a bed of textures,'' where a few discrete sound events are superimposed upon a stationary acoustic background~\cite{nelken2013}.
Such events are not well-characterised by summarizing short-term features, but are better described by large-scale temporal evolution of auditory scenes.
The latter approach should therefore prove more fruitful in measuring auditory scene similarity.

@@

Many algorithms in pattern recognition, including nearest neighbor classifiers and support vector machines, tend to work best when all features follow a standard normal distribution across all training instances~\cite{Hsu2003}.
Yet the distribution of the scattering coefficients is skewed towards larger values. We can reduce this skewness by applying a pointwise concave transformation to all coefficients. In particular, we find that the logarithm performs particularly well in this respect.
Figure~\ref{fig:histograms} shows the distribution of an arbitrarily chosen scattering coefficient over the DCASE 2013 dataset, before and after logarithmic compression.

@@

The experiments in this paper are carried out on the publicly available DCASE 2013 dataset~\cite{7100934}.
Although the dataset was constructed for the task of acoustic scene classification, where the goal is to correctly assign the class of a given recording, we can use the same recordings and class labels for the task of acoustic scene similarity retrieval.
The dataset consists of two parts, a public and a private subset, each made up of $100$ acoustic scene recordings sampled at $44100\,\mathrm{Hz}$ and $30$ seconds in duration. The dataset is evenly divided into $10$ acoustic scene classes: \emph{bus}, \emph{busy street}, \emph{office}, \emph{open air market}, \emph{park}, \emph{quiet street}, \emph{restaurant}, \emph{supermarket}, \emph{tube}, and \emph{tube station}.
The recordings were made by three different recordists at a wide variety of locations in the Greater London area over a period of several months.
In order to avoid any correlation between recording conditions and label distribution, all recordings were carried out under moderate weather conditions, at varying times of day and week, and each recordist recorded each scene type. As a result, the dataset enjoys significant intra-class diversity while remaining of manageable size, making it suitable for evaluation of algorithmic design choices~\cite{lagrange:hal-01082501}.

@@

The scaling parameter $q$ of the RBF kernels (see Eq.~\ref{eq:kc}) is set to $10\%$ of the number of data points to cluster.  As the number of Gaussians for the BoF approach, the numbers of clusters $M$ controls the level of abstraction. For each method, unless otherwise stated, the parameter $M$ is set to $8$. It thus allows $8$ different types of observations to be modeled, which seems reasonable given the duration of the scene (30 seconds). Note that this is the only free parameter in the proposed method. However, except for $\emph{RbQ-a}$, the results are not very sensitive to the choice of $M$, as long as it is large enough to characterize the seen. A numerical demonstration is provided at the end of the next section.
